Secure Composition of Robust and Optimising Compilers

### 2023-07-08 popl/main.tex
Sketch intro

\hltt{Context}
\hltt{Problem}

\begin{lstlisting}[language=c,basicstyle=\small\ttfamily, caption=C]
int strncpy(int n, int *x, int *y) {
  for(int i = 0; x[i] != '\0' && i < n; ++i) {
    y[i] = x[i];
  }
}
\end{lstlisting}

\BrText{Solution}{
% keyword: pass
This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
Then, this paper defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
% keyword- end to end, composition
Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by the framework of this paper.
The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
\MPin{
	last line: not the best.
}
}

% \BrText{Validation}{
In summary, this paper makes the following contributions:

@@

  \item The key contributions are formalized in the Coq proof assistant.
        The practically relevant parts of the formalization are \emph{extractable}.
        That is, one can translate the Coq code to Ocaml and run the compilation chain from the previous contribution.
        Finally, this paper indicates with \CoqSymbol whenever definitions or proofs are done in Coq.
\end{itemize}
% }

>>>>

\gls{ms} is a security property of programs.
Without memory safety, private data can easily leak or corrupt~\cite{lemay2021ccc}.
Because of this, \gls{ms} is a necessity for programs to be \gls{cct}, which is an important security property for secrecy.
Programs that attain \gls{cct} are secure with respect to timing attacks~\cite{kocher1996timing}.
Timing attacks may expose secrets from a simple performance analysis.
These attacks have become more and more relevant since \citeauthor{kocher1996timing}'s paper, especially given the era of cloud--computing~\cite{aviram2010cloudtime,kumar2019cloudsecsurvey}, where programs share the same hardware but are supposed to run in a sandboxed environment that does not leak secrets.
Unfortunately, several attacks within such a cloud environment have been demonstrated as viable strategies to extract secrets~\cite{mehmet2015getoff,flowers2022zeroday,atya2019catchme,venkatanathan2015placevul}.

Even though mitigations for such attacks exist~\cite{bond2017vale,almeida2017jasmin}, it has been shown in literature that some compilers may optimize these mitigations away~\cite{barthe2018sec}.
This is of course not what the programmer expects and necessitates the need for a {\em secure} compiler that ensures security properties hold even when the program runs in a hostile environment, i.e., interacts with other programs by any means.
Because of this hostile environment, secure compilers attain security by relying on certain protection mechanisms, such as ... and ...\MK{cits}.
The aim of this paper is to devise the theoretical foundation to be able to build a secure compiler that is capable of dealing with the following example.
Consider the program bolow which copies a \texttt{C}-style string \texttt{x} into \texttt{y} up to a length of \texttt{n}:

\begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
int strncpy(char n, char *x, char *y) {
  for(int i = 0; x[i] != '\0' && i < n; ++i) {
    y[i] = x[i];
  }
}
\end{lstlisting}

For sake of simplicity it is assumed that the buffer \texttt{y} is at least as large as \texttt{n} as well as \texttt{x} terminated by \texttt{'\textbackslash 0'}.
Even under these assumptions, there is a subtle \gls{ms} bug in the program which can be exploited to read sensitive data:
The bounds check, i.e., \texttt{i < n}, should happen prior to the access to memory location \texttt{x[i]}.
Because otherwise, if \texttt{n} is larger than the length of the array object \texttt{x} points to, the memory location past this array object is being read, which is an out--of--bounds memory access.
Unfortunately, even though flipping the conditions does fix the \gls{ms} issue, \gls{cct} is violated, so the program is susceptible to timing attacks.
This is because, the loop can terminate early: As soon as the string--terminating character \texttt{'\textbackslash 0'} is encountered, the loop exits.
Thus, program execution time is proportional to the length of the array object \texttt{x}.
The fix commonly employed in such a case is to rewrite the program to run a fixed number of iterations regardless of the length of \texttt{x} or \texttt{y}.
To this end, there must be an upper bound for the length of \texttt{x} and \texttt{y} and then the loop can just iterate until the induction variable \texttt{i} hits this upper bound.

These fixes alone are not enough to make this program robust, i.e., not violating either \gls{ms} or \gls{cct} in a malicious environment.
Consider a program context that calls \texttt{strncpy} with a very large \texttt{n}, some uninitialised memory \texttt{x}, and \texttt{y} whatever.
It is easy to see that such a function invocation leads to a series of reads from uninitialised memory, an immediate \gls{ms} issue.
To mitigate these situtations, it is important that \texttt{strncpy} ensures that the provided arguments satisfy the assumptions being made.
After all, the environment cannot be trusted and potentially behaves in a hostile manner.

\MKin{
  show final implementation?
}
%\begin{lstlisting}[language=c,basicstyle=\small\ttfamily]
%int strncpy(char n, char *x, char *y) {
%  for(int i = 0; x[i] != '\0' && i < n; ++i) {
%    y[i] = x[i];
%  }
%}
%\end{lstlisting}

As stated earlier, this work wants to establish the theoretical foundation to ensure the secure compilation of programs like \texttt{strncpy} to satisfy properties \gls{ms} and \gls{cct} robustly.
To this end, this paper chooses the robust preservation framework, which is a model of secure compilation that is well--established~\cite{abate2019jour,abate2021extacc,patrignani2021rsc}. % and has a stronger attacker model than a passive attacker whose only means of interacting with a program are timing analyses and so on.
Unfortunately, this framework of secure compilers does not support modularity: One has to make a secure compiler for all possible properties {\em or} for a combination of \gls{ms} and \gls{cct}.

\MKin{
blablabla continue blablabla
}

% keyword: pass
This paper introduces a framework for reasoning about the composition of secure and optimising compiler passes and it showcases the power of this framework by instantiating it on a multi-pass compilation chain.
To this end, this paper first discusses how to compose security properties, such as temporal and spatial memory safety into general memory safety, and cryptographic constant time.
Then, this paper defines several secure compiler passes, where each is either preserving a different security property (e.g., temporal or spatial memory safety) or performing a security-preserving optimisation, (e.g., applying constant-folding or dead-code elimination).
% keyword- end to end, composition
Finally, this paper shows how to compose these secure compiler passes into a multi-pass compilation chain which is proven to provide end-to-end preservation of general memory safety and cryptographic constant time.
Crucially, this end-to-end security preservation is obtained by having each compilation pass (i) preserve a sub-part of the overall security property individually and then (ii) compose each preserved security property as dictated by the framework of this paper.
The results showcase how the framework allows the kind of formal security reasoning that compiler writers already want (and do), obtaining precise, compositional security reasoning while providing minimal (and modular) proof effort.
\MPin{
	last line: not the best.
}

In summary, this paper makes the following contributions:

@@

  \item The key contributions are formalized in the Coq proof assistant.
        The practically relevant parts of the formalization are \emph{extractable}.
        That is, one can translate the Coq code to Ocaml and run the compilation chain from the previous contribution.
        Finally, this paper indicates with \CoqSymbol whenever definitions or proofs are done in Coq.
\end{itemize}
